Cassini State Capture

### 2019-10-15 writeups/Su_disk_disp.tex
probabilities are now being overplotted, as well as III -> III

\section{Equations}\label{s:eq}

Denote $\vec{s}$ spin of planet, $\vec{l}$ angular momentum of planet, and
$\vec{l}_d$ angular momentum of the surrounding disc. We approximate $S \ll L
\ll L_d$, so $\vec{l}_d$ is approximately constant. Much of this treatment
borrows from \citealp{anderson2018teeter} and \citealp{millholland_disk}. We
consider precession equations
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}}\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{ld}\p{\hat{l} \cdot \hat{l}_d}
        \p{\hat{l} \times \hat{l}_d},\label{eq:dldt}\\
\end{align}
where
\begin{align}
    \omega_{sl} &\equiv \frac{3k_{qp}}{2k_p} \frac{M_\star}{m_p}
        \p{\frac{R_p}{a_p}}^3 \Omega_p,\label{eq:wsl}\\
    \omega_{ld} &\equiv \frac{3M_d}{4M_\star}\p{\frac{a_p}{r_d}}^3 n
        .\label{eq:wld}
\end{align}
In \autoref{eq:wsl}, $k_{qp}$ and $k_p$ are related to the planet's quadrupole
moment and moment of inertia respectively \citep[see][]{lai2018}, $M_\star$ is
the mass of the host star, and $M_p, R_p, a_p, \Omega_p$ are the mass, radius,
semi-major axis and angular spin frequency of the planet respectively.
\autoref{eq:wld} is written for a disk of mass $M_d$ located at characteristic
distance $r_d$ from the host star \citep[see][for a power-law disk
profile]{millholland_disk}, $n \equiv \sqrt{GM_\star/a_p^3}$ is the planet's
orbital mean motion. The three angles formed by the three vectors are notated
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{sd} &\equiv \arccos \hat{s} \cdot \hat{l}_d,\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_d.
\end{align}

We go to the corotating frame with $\hat{l}$ fixed such that $\hat{z} =
\hat{l}$ ($\theta$ is now conveniently also the polar component of $\hat{s}$).
The evolution of $\hat{s}$ in this corotating frame is governed by:
\begin{equation}
    \rd{\hat{s}}{t} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \abs{g}\p{\hat{s} \times \hat{l}_d}.
\end{equation}
$\alpha = \omega_{sl} > 0, g \equiv -\omega_{ld}\cos I < 0$ follow traditional
sign conventions.

We divide through by $\alpha$ and define parameter $\eta$ and nondimensionalized
time $\tau$
\begin{align}
    \eta &\equiv \frac{\abs{g}}{\alpha}\label{eq:eta},\\
    \tau &\equiv \alpha t.
\end{align}
The redimensionalized equation of motion is then
\begin{equation}
    \rd{\hat{s}}{\tau} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_d}. \label{eq:dsdt_base}
\end{equation}

Following~\cite{millholland_disk}, we allow $g$ to vary in time owing to a
disk with decaying mass
\begin{equation}
    M_d(t) = M_d(0)e^{-t/t_d}.
\end{equation}
As $\alpha$ is held constant, this equates to $\eta$ decaying as $\rd{\eta}{t} =
-\frac{\eta}{t_d}$ for some decay time $t_\eta$. We nondimensionalize by
defining $\epsilon \equiv \frac{1}{\alpha t_\eta}$, so
\begin{equation}
    \rd{\eta}{\tau} = -\epsilon \eta.\label{eq:deta_dt}
\end{equation}
Equations \autoref{eq:dsdt_base} and \autoref{eq:deta_dt} constitute our system
of study.

Finally, the Hamiltonian corresponding to the equation of motion
\autoref{eq:dsdt_base} is
\begin{equation}
    \mathcal{H}\p{\theta, \phi} = -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
        + \eta \p{\hat{s} \cdot \hat{l}_d}.\label{eq:H}
\end{equation}
Here, we adopt a spherical coordinate system where $\hat{l} = \hat{z}$ and
$\theta, \phi$ are the polar and azimuthal angle of $\hat{s}$. By convention, we
choose $\hat{l}_z$ at coordinates $\theta = I, \phi = \pi$, in which case the
Hamiltonian can be written
\begin{equation}
    \mathcal{H} = -\frac{1}{2}\cos^2\theta
        + \eta \p*{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{equation}

Note that $\p{\cos \theta, \phi}$ form a canonically conjugate pair of variables
describing $\hat{s}$. The equations of motion for these variables can be derived
by taking derivatives of $\mathcal{H}$ given by \autoref{eq:H} and agree with
\autoref{eq:dsdt_base}:
\begin{subequations}\label{se:H_eom}
    \begin{align}
        \pd{\phi}{t} = \pd{\mathcal{H}}{(\cos\theta)}
            &= -\cos\theta + \eta\p{\cos I + \sin I \cot \theta \cos \phi},\\
        \pd{(\cos \theta)}{t} = -\pd{\mathcal{H}}{\phi}
            &= -\eta \sin I \sin \theta \sin \phi.
    \end{align}
\end{subequations}
While these equations are numerically stiff owing to the $\cot\theta$ term, they
are the most intuitive description for analytical work.

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CSs). There are either two or four solutions, depending on the values
of $\eta, I$. Following standard convention and nomenclature (see
\autoref{fig:cs_locs}), CSs 1, 3, 4 have $\phi = 0, \theta < 0$ and CS2 has
$\phi = \pi, \theta > 0$. Of these, CSs 1, 2, 3 are stable while CS4 is
unstable.

When $\eta < \eta_c$, where
\begin{equation}
    \eta_c \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},
\end{equation}
all four CSs exist, and when $\eta > \eta_c$, only CSs 2, 3 exist
\citep{henrard1987,ward2004I}. The CS obliquities as a function of $\eta$ are
provided in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{$\theta$ values of the Cassini states over time, following the same
    coloring scheme as \autoref{fig:eq_1contours}. Note that $\theta \in [-\pi,
    \pi]$ is the traditional definition of the polar angle
    \citep[see e.g.][]{colombo1966,peale1969,henrard1987}.}\label{fig:cs_locs}
\end{figure}

@@

When $\eta < \eta_c$, CS4 exists and is a saddle point. All trajectories are
periodic with finite period except two critical trajectories asymptotic in the
past and future to CS4. Together, these two critical trajectories are referred
to as the \emph{separatrix} and divide phase space into three zones. The
separatrix, three zones, and their relations to the CSs are most clearly shown
via a contour plot of $\mathcal{H}\p{\phi, \cos \theta}$ in
\autoref{fig:eq_1contours}.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$, where warmer
    colors denote more positive values. The black solid line is the separatrix,
    which only exists for $\eta < \eta_c$. The three zones, divided by the
    separatrix, are labeled. The interior of the separatrix, shaded in grey, is
    formally only defined for $\eta < \eta_c$, but we may identify the phase
    space trajectories that flow into zone II when evolved forward in time; this
    is the shaded region in the top left plot, bounded by the dotted black
    line.}\label{fig:eq_1contours}
\end{figure*}

As trajectories of constant $\eta$ evolve along contours of constant
$\mathcal{H}$, it can be observed that trajectories in zone II librate about CS2
while those in zones I/III circulate.

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{align*}
    z_0 &= \eta\cos I, &
    \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
    \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
        \chi^2 \cos^2\theta_4 + 1},&
    T &= 2\chi \frac{\cos \theta_4}{
        \chi^2 \cos^2\theta_4 - 1},\\
\end{align*}
the areas for $\eta < \eta_c$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_I &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_c$, but a natural extension
exists: evolve an initial point $p$ under adiabatic decrease of $\eta > \eta_c$
until the separatrix appears at $\eta = \eta_c$, then identify $p$ with the zone
it is in at $\eta_c$. Since phase space area is conserved under adiabatic
evolution, this implies $A_i\p{\eta > \eta_c} = A_i(\eta_c)$.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional $A_{i}(\eta)$ as given by
    \autoref{se:area_ward}. Dotted lines correspond to small $\eta$
    approximations discussed in \autoref{ss:app_transition}. Dashed lines denote
    effective values of $A_{II}, A_{III}$ for $\eta > \eta_c$, denoting the
    area of points that would flow into either area under adiabatic decrease of
    $\eta$ to below $\eta_c$.}\label{fig:eq_areas}
\end{figure}

@@

As $\eta$ decreases, the trajectory generally encounters the separatrix at some
value $\eta_\star$ depending on its initial condition. If $\eta$ changes
sufficiently little during this separatrix-crossing orbit, the evolution of the
system is considered \emph{adiabatic}. In the truly adiabatic limit, $\epsilon
\to 0$; we use $\epsilon = 3 \times 10^{-4}$ unless otherwise noted.

One necessary criterion for adiabaticity is that the separatrix crossing
timescale be longer than the the resonant libration period \citep{ward2004I}.
The resonant libration period is just the libration period about CS2, which can
be found by linearizing \autoref{se:H_eom} about $\phi \approx \pi, \theta
\approx \theta_2$. The resulting criterion is
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}}_{cross} = \epsilon &\lesssim
            \frac{2\pi}{T_{lib}},\\
        &\lesssim \sqrt{\eta_\star\sin\theta \sin I
            \p{1 + \eta_\star \frac{\sin I}{\sin^3\theta}}}.\label{eq:ad_constr}
\end{align}
Note that this is a necessary but not sufficient criterion: the libration period
is longer for trajectories nearer the separatrix and diverges at the separatrix
for constant $\eta$. This formula also is in slight disagreement with that
presented in \citealt{millholland_disk}

@@

@@

Next, we will consider the evolution of a system with arbitrary $\theta_{sd, i}$
starting in the $\eta \gg 1$ regime. By evolving such systems until $\eta \ll
1$, we can measure the final $\theta_f$ values and make comparisons with those
obtained via the dynamical tracks above. Observe that when $\eta \gg 1$,
trajectories librate about $\hat{l}_d$, or equivalently CS2, with approximately
constant $\theta_{sd}$, meaning they enclose phase space area
\begin{equation}
    A_i = 2\pi\p{1 - \cos \theta_{sd, i}}.\label{eq:ai_qsd}
\end{equation}
Thus, for a given $\theta_{sd, i}$, we know immediately which tracks are
accessible and with what probability following the results of
\autoref{ss:zone_transitions}.

We run simulations over a ring of initial conditions at each $\theta_{sd, i}$
and measure the final $\theta_{f}$ values. Since we start with finite $\eta$, we
use the angular distance to CS2 as a proxy for $\theta_{sd, i}$, as it is the
correct quantity to use to make \autoref{eq:ai_qsd} hold exactly. We use
\autoref{eq:ai_qsd} to predict the final obliquities $\theta_f$ according to
each of the dynamical tracks. The agreement of these predictions for $I =
5^\circ, 10^\circ, 20^\circ$ can be seen in \autoref{fig:ad_ensemble},
\autoref{fig:3_ensemble_10_35}, and \autoref{fig:3_ensemble_20_35}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{f}\p{\theta_{sd, i}}$, overlaid with semi-analytic
    predictions of the $\theta_{ f}$ for each of the four nontrivial
    dynamical tracks in colored lines. Bottom: Semi-analytic probabilities of
    each of the dynamical tracks for each $\theta_{sd, i}$. The five regimes of
    $\theta_{sd, i}$ values correspond to the five regimes of $A_i$ discussed in
    \autoref{ss:zone_transitions} when using \autoref{eq:ai_qsd} to relate
    $A_i\p{\theta_{sd, i}}$. }\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_20_35}
\end{figure}

>>>>

\section{Theory}\label{s:eq}

\subsection{Equations of Motion}

Denote $\hat{s}$ spin of planet, $\hat{l}$ angular momentum of planet, and
$\hat{l}_d$ angular momentum of the surrounding disc. We approximate $S \ll L
\ll L_d$, so $\hat{l}_d$ is approximately constant and $\hat{l}$ experiences
negligible backreaction torque from $\hat{s}$. Much of this treatment borrows
from \citealp{anderson2018teeter} and \citealp{millholland_disk}. We consider
precession equations
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}}\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{ld}\p{\hat{l} \cdot \hat{l}_d}
        \p{\hat{l} \times \hat{l}_d}\label{eq:dldt}.
\end{align}
where
\begin{align}
    \omega_{sl} &\equiv \frac{3k_{qp}}{2k_p} \frac{M_\star}{m_p}
        \p{\frac{R_p}{a_p}}^3 \Omega_p,\label{eq:wsl}\\
    \omega_{ld} &\equiv \frac{3M_d}{4M_\star}\p{\frac{a_p}{r_d}}^3 n
        .\label{eq:wld}
\end{align}
In \autoref{eq:wsl}, $k_{qp}$ and $k_p$ are related to the planet's quadrupole
moment and moment of inertia respectively \citep[see][]{lai2018}, $M_\star$ is
the mass of the host star, and $M_p, R_p, a_p, \Omega_p$ are the mass, radius,
semi-major axis and angular spin frequency of the planet respectively.
\autoref{eq:wld} is written for a disk of mass $M_d$ located at characteristic
distance $r_d$ from the host star \citep[see][for a power-law disk
profile]{millholland_disk}, $n \equiv \sqrt{GM_\star/a_p^3}$ is the planet's
orbital mean motion. The three angles formed by the three vectors are notated
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{sd} &\equiv \arccos \hat{s} \cdot \hat{l}_d,\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_d.
\end{align}

Let us transform to frame rotating about $\hat{l}_p$ with frequency $g \equiv
-\omega_{ld}\cos I < 0$, where $\hat{l}$ fixed. In this frame, $\hat{l}_p$
also remains fixed and $\hat{s}$ evolves as:
\begin{equation}
    \rd{\hat{s}}{t} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        + g\p{\hat{s} \times \hat{l}_d},
\end{equation}
where $\alpha \equiv \omega_{sl} > 0$ follows standard notation. By defining
$\eta$ and nondimensionalized time $\tau$ as
\begin{align}
    \eta &\equiv \frac{\abs{g}}{\alpha}\label{eq:eta},\\
    \tau &\equiv \alpha t,
\end{align}
we can non-dimensionalize the equation of motion as
\begin{equation}
    \rd{\hat{s}}{\tau} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_d}. \label{eq:dsdt_base}
\end{equation}

Following~\cite{millholland_disk}, we allow $g$ to vary in time owing to a
disk with decaying mass
\begin{equation}
    M_d(t) = M_d(0)e^{-t/t_d}.
\end{equation}
As $\alpha$ is held constant, this equates to $\eta$ decaying as $\rd{\eta}{t} =
-\frac{\eta}{t_d}$ for some decay time $t_\eta$. We nondimensionalize by
defining $\epsilon \equiv \frac{1}{\alpha t_\eta}$, so
\begin{equation}
    \rd{\eta}{\tau} = -\epsilon \eta.\label{eq:deta_dt}
\end{equation}
\autoref{eq:dsdt_base} and \autoref{eq:deta_dt} together constitute our system
of study.

For analytical work, we adopt spherical coordinate system where $\hat{l} =
\hat{z}$ and $\theta, \phi$ are the polar and azimuthal angle of $\hat{s}$. By
convention, we choose $\hat{l}_z$ at coordinates $\theta = I, \phi = \pi$.
Note that $\p{\cos \theta, \phi}$ form a canonically conjugate pair of variables
describing $\hat{s}$. To derive equations of motion for these canonical
coordinates, we introduce Hamiltonian
\begin{align}
    \mathcal{H}\p{\theta, \phi} &= -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_d}.\label{eq:H}\\
        &= -\frac{1}{2}\cos^2\theta
            + \eta \p{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{align}

The equations of motion follow by taking derivatives of $\mathcal{H}$ and agree
with \autoref{eq:dsdt_base}:
\begin{subequations}\label{se:H_eom}
    \begin{align}
        \pd{\phi}{t} = \pd{\mathcal{H}}{(\cos\theta)}
            &= -\cos\theta + \eta\p{\cos I + \sin I \cot \theta \cos \phi},\\
        \pd{(\cos \theta)}{t} = -\pd{\mathcal{H}}{\phi}
            &= -\eta \sin I \sin \theta \sin \phi.
    \end{align}
\end{subequations}
While these equations are numerically stiff owing to the $\cot\theta$ term, they
are the most intuitive description for analytical work.

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CSs). There are either two or four solutions, depending on the values
of $\eta, I$. Following standard convention and nomenclature (see
\autoref{fig:cs_locs}), CSs 1, 3, 4 have $\phi = 0, \theta < 0$ and CS2 has
$\phi = \pi, \theta > 0$. Of these, CSs 1, 2, 3 are stable while CS4 is unstable
(this follows from linearization of \autoref{se:H_eom}). When $\eta < \eta_c$,
where
\begin{equation}
    \eta_c \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},
\end{equation}
all four CSs exist, and when $\eta > \eta_c$, only CSs 2, 3 exist
\citep{henrard1987,ward2004I}. The CS obliquities as a function of $\eta$ are
shown in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{$\theta$ values of the Cassini states over time, following the same
    coloring scheme as \autoref{fig:eq_1contours}. Note that $\theta \in [-\pi,
    \pi]$ is the traditional definition of the polar angle
    \citep[see e.g.][]{colombo1966,peale1969,henrard1987}.}\label{fig:cs_locs}
\end{figure}

@@

When $\eta < \eta_c$, CS4 exists and is a saddle point. The two trajectories
originating and ending at CS4 are the only two infinite-period orbits in phase
space. Together, these two critical trajectories are referred to as the
\emph{separatrix} and divide phase space into three zones. The separatrix, three
zones, and their relations to the CSs are shown in \autoref{fig:eq_1contours}.
Note that trajectories in zone II librate about CS2 while those in zones I/III
circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$, where warmer
    colors denote more positive values. The black solid line is the separatrix,
    which only exists for $\eta < \eta_c$. The three zones, divided by the
    separatrix, are labeled. The interior of the separatrix, shaded in grey, is
    formally only defined for $\eta < \eta_c$, but we may identify the phase
    space trajectories that flow into zone II when evolved forward in time; this
    is the shaded region in the top left plot, bounded by the dotted black
    line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{align*}
    z_0 &= \eta\cos I, &
    \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
    \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
        \chi^2 \cos^2\theta_4 + 1},&
    T &= 2\chi \frac{\cos \theta_4}{
        \chi^2 \cos^2\theta_4 - 1},\\
\end{align*}
the areas for $\eta < \eta_c$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_I &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_c$, but a natural extension
exists: evolve an initial point $p$ under adiabatic decrease of $\eta > \eta_c$
until the separatrix appears at $\eta = \eta_c$, then identify $p$ with the zone
it is in at $\eta_c$. Since phase space area is conserved under adiabatic
evolution, this implies $A_i\p{\eta > \eta_c} = A_i(\eta_c)$. This extension is
also reflected in the dashed boundary in \autoref{fig:eq_1contours}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional $A_{i}(\eta)$ as given by
    \autoref{se:area_ward}. Dotted lines correspond to small $\eta$
    approximations discussed in \autoref{ss:app_transition}. Dashed lines denote
    effective values of $A_{II}, A_{III}$ for $\eta > \eta_c$, denoting the
    area of points that would flow into either area under adiabatic decrease of
    $\eta$ to below $\eta_c$.}\label{fig:eq_areas}
\end{figure}

@@

A perturbation to a Hamiltonian system is considered \emph{adiabatic} if it is
sufficiently slow compared to all other relevant time scales in the problem.
In the current problem, the slowest time scale is the \emph{separatrix crossing
time scale}: as $\eta$ evolves, a trajectory may encounter the separatrix at
some value $\eta_\star$ which varies depending on its initial condition. If
$\eta$ changes sufficiently little during the separatrix-crossing orbit, the
evolution of the system may be considered adiabatic. The truly adiabatic limit
corresponds to $\epsilon \to 0$; we take $\epsilon = 3 \times 10^{-4}$ unless
otherwise noted.

One criterion for adiabaticity is that the separatrix crossing timescale be
longer than the the resonant libration period \citep{ward2004I}. The resonant
libration period is just the libration period about CS2, which can be found by
linearizing \autoref{se:H_eom} about CS2. This results in constraint
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}}_{cross} = \epsilon &\lesssim
            \frac{1}{T_{lib}},\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta_\star\sin\theta \sin I
            \p{1 + \eta_\star \frac{\sin I}{\sin^3\theta}}}.\label{eq:ad_constr}
\end{align}
This formula differs from that presented in \citealt{millholland_disk}.

We consider the evolution of a system with arbitrary initial $\theta_{sd, i}
\equiv \theta_{sd}\p{t = 0}$ and $\eta_i \gg 1$. We are interested in the
distribution of final $\theta_f$ values after $\eta$ decreases to $\eta_f \ll 1$
(after the disk has dissipated to negligible mass). While it is likely that
$\theta_{sd, i} \approx 0$ as the planet forms in the disk, we explore all
$\theta_{sd, i} \in [0, \pi]$ for completeness.

We evenly sample $101$ values of $\theta_{sd, i}$, and for each $\theta_{sd, i}$
value, we pick $101$ $\hat{s}$ initial conditions on the ring fixed by
$\theta_{sd, i}$. To be concrete, we choose $\eta_i = 10\eta_c$. For each
initial condition, we evolve Equations~\ref{eq:dsdt_base} and~\ref{eq:deta_dt}
until $\eta \leq 10^{-5}$, where $\hat{s}$ precesses about $\hat{l}$ with
uniform final obliquity $\theta_f$. This transfer function
$\theta_f\p{\theta_{sd, i}}$ is our primary result, and is shown for $I =
5^\circ, 10^\circ, 20^\circ$ in \autoref{fig:ad_ensemble},
\autoref{fig:3_ensemble_10_35}, and \autoref{fig:3_ensemble_20_35} respectively.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{f}\p{\theta_{sd, i}}$, overlaid with semi-analytic
    predictions of the $\theta_{ f}$ for each of the four nontrivial
    dynamical tracks in colored lines. Bottom: Semi-analytic probabilities of
    each of the dynamical tracks for each $\theta_{sd, i}$. The five regimes of
    $\theta_{sd, i}$ values correspond to the five regimes of $A_i$ discussed in
    \autoref{ss:zone_transitions} when using \autoref{eq:ai_qsd} to relate
    $A_i\p{\theta_{sd, i}}$. }\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$.}\label{fig:3_ensemble_20_35}
\end{figure}

@@

Observe that when $\eta \gg 1$, trajectories librate about $\hat{l}_d$, or
equivalently CS2, with approximately constant $\theta_{sd}$, meaning they
enclose phase space area
\begin{equation}
    A_i = 2\pi\p{1 - \cos \theta_{sd, i}}.\label{eq:ai_qsd}
\end{equation}
In our simulations, we start with finite $\eta$, so we use the angular distance
to CS2 instead of $\theta_{sd, i}$ above, as the trajectory initially librates
about CS2 and not $\hat{l}_p$.

@@
